Fisheries management when quotas are costly to change

### 2014-07-25 manuscripts/manuscript.Rmd
updated presentation, minor tweaks to manuscript from Paul feedback

---
title: Optimal management of a stochastically varying population when policy adjustment is costly
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviera
    affiliation: UTK
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul Armsworth 
    affiliation: UTK

@@

@@

```{r Figure_1, fig.cap="Quota on Blue Fin Tuna Catch", fig.width=4, fig.height=4, dev=c("pdf", "svg", "png")}
tuna <- read.csv("components/data/tuna.csv")
ggplot(tuna, aes(year, quota)) + geom_point()
```

@@

As the example in Figure 1 makes clear, management policies will rarely be
as responsive as this constant escapement policy assumes. In this paper,
we explore conditions under which this lack of responsiveness may be a
rationale response on the part of managers. Specifically, we consider a
case where managers seek to balance the benefits in terms of increased
profits from fishing from more finely tracking recruitment variations
with the growing costs associated with adjusting policies quickly to do
so. The policy adjustment costs involved could reflect pure administrative
transaction costs or preferences held by fishermen, fish processing plants
or other stakeholders, for less variable quotas. In seeking to account
for these policy adjustment costs, we recognize that we do not know just
what functional form they should take and that it will require a lot of
empirical work to estimate that. Therefore, we scope three candidate
functional forms that represent different assumptions about how these
costs operate to determine whether the results we obtain are sensitive
to such differences.

@@

```{r Figure_2, fig.cap="Expected net present value by functional form of penalty. Horizontal axis shows the coefficient $c_i$ governing the magnitude of the policy cost, while vertical axis shows fraction of the maximum expected net present value dissipated by the policy cost, $\\tfrac{NPV_0 - NPV_i}{NPV_0}$.  The horizontal line indicates a value of the stock that is reduced by `r reduction * 100`% from the maximum expected value in the absence of policy adjustment costs, $NPV_0({\\bf h_0^*})$. Selecting the coefficient $c_i$ corresponding to this value in each functional form allows us to make consistent comparisons across the different functional forms of policy costs.", dependson="quadcosts", fig.width=4, fig.height=3, cache = fig.cache, dev=c("pdf", "svg", "png")}

@@

ggplot(fees, aes(c2, (npv0-value)/npv0, lty=variable)) + 
  geom_line() + 
  geom_hline(aes(yintercept=reduction), linetype=4) + 
  xlab("Penalty coefficient") +
  ylab("Reduction in net present value") + 
  scale_linetype_discrete(labels=relabel)

@@

```{r Figure_3, dependson="tidy", fig.cap="Example realization of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=3, fig.width=5, cache = fig.cache, dev=c("pdf", "svg", "png")}

@@

The comparisons shown in Figure 3 are for one realization (i.e. a
particular sequence of random number draws representing environmental
variability) and are made for a particular magnitude of policy
adjustment costs equal to `r reduction`% of the maximum expected net present value 
in the base case ($NPV_0({\bf h_0^*})$). Figure 4 generalizes to show
how optimal harvest levels and corresponding optimal stock sizes are
affected by the magnitude of the costs of policy adjustment. This time
we show summary statistics across 50 replicate sets of environmental
conditions. We show the impact of policy adjustment costs on the
variance of both harvests and stock sizes through time; on the
autocorrelation of harvests and stock sizes through time. <!-- NOTE Are we showing cross-correlations or not? --> <!-- See figure 4, probably not-->

For example, when small policy adjustment costs cost little but large
adjustments are expensive ($\Pi_2$), we see the smoothing signal that we
might have expected (see also @Ludwig1980 for this particular case). As
cost penalties increase in severity (i.e., $c_2$ increases), the
variance in quotas through time decreases and the autocorrelation of
quotas through time increases. This smoothing of the harvest quotas
through time has knock-on effects for the corresponding optimal stock
sizes, which also become more autocorrelated as the scaling on policy
adjustment costs ($c_2$) is increased. The autocorrelation in stock
sizes here reflects the lag involved in a more modest increase in
harvests taking longer to eliminate the signal of strong recruitment
events, when compared to a more immediately responsive constant
escapement policy.

@@

Finally, the case where policy adjustment costs scale linearly with the
size of the adjustment ($\Pi_1$ in Eqn. 4) appear to be something of a
middle of the road strategy, in that increasing the severity of policy
adjustment costs (increasing $c_1$ has little effect on the variance or
autocorrelation of optimal harvest rates or stock sizes. To reveal the
particular impact of policy adjustment costs of this type requires a
more targeted summary statistic. Specifically, for each run we
calculated the frequency with which the optimal policy involved
maintaining a positive quota across multiple time steps unaltered. This
type of policy is arguably the most commonly observed behavior in TAC
management, but is one that is almost never observed to be part of the
optimal management strategy in the basic model without policy adjustment
costs (Eqn. 3) or when optimizing against $\Pi_2$ or $\Pi_3$.

@@

While still not common for the particular parameter combinations we examine,
we find that positive unaltered quotas through time are much more likely
to occur when optimizing against $\Pi_1$ where policy adjustment costs
scale linearly with the size of proposed policy changes: in the 100 replicates
of data presented above the harvest policy is strictly positive and identical in consecutive intervals
only `r no_shifts[["L2"]] * 100`% of the time for quadratic costs $\Pi_2$, compared with
`r no_shifts[["L1"]] * 100` with linear costs $\Pi_1$ and `r no_shifts[["fixed"]]*100`%
for fixed costs $\Pi_3$. Moreover, these occurrences increase in frequency as the severity of
these costs ($c_1$) increases.

```{r Figure_4, fig.cap="Variance and autocorrelation of stock dynamics and harvest dynamics, as a function of the penalty coefficient.  Adjustment penalty is calculated as percentage of the maximum expected NPV in the basic case with no policy adjustment costs.", fig.height=8, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png")}
ggplot(subset(figure4_df, statistic != "cross.correlation"), 
       aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn))  +
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  coord_cartesian(xlim = c(0, .3)) + 
  facet_wrap(~ timeseries + statistic, scale="free_y") + 
  scale_color_discrete(labels=relabel, name="Penalty function") + 
  scale_fill_discrete(labels=relabel, name="Penalty function") +
  xlab("Penalty size (as fraction of NPV0)")

write.csv(figure4_df, "components/data/figure4.csv")
```

@@

```{r Figure_5, fig.cap="Distribution of profits from fishing and costs paid to adjustment.  Distributions are shown when managing a system in which the adjustment costs present are determined by the model indicated, and compared to the same costs and profits of managing the same system using a Reed model, which ignores adjustment costs, and thus tends to pay much higher fees for adjustment.  In contrast, the difference in profit after accounting for costs is only marginally less than the value derived by managing the population to optimize profits when ignoring adjustment costs (Reed model).", fig.width=6, fig.height=5, dependson="histogram-calc", dev=c("pdf", "svg", "png")}

@@

```{r Figure_6, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue), shown by penalty function over increasing magnitudes of cost.  All values are relative to the cost-free adjustment optimum, NPV0.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. The severity of the latter mistake increases sharply as adjustment costs make a larger fraction of the total costs.", dev=c("pdf", "svg", "png")}

@@

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure 7 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10%
less than optimal free. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

For fixed costs ($\Pi_3$), the reverse is true.  Because fixed costs increase the
volatility of the optimal solution while $\Pi_2$ costs decrease it, simply
ignoring the cost of adjustment all together results in a higher value
than either assuming costs are quadratic ($\Pi_2$) when they are fixed ($\Pi_3$),
or assuming they are fixed when they are quadratic.

Nevertheless, assuming quadratic costs is always the most conservative
option -- scenarios in which true costs are quadratic (regardless of what
cost structure is assumed by the policy) do significantly worse than
those that are not (i.e. all cases that assume quadratic but actually
have linear or fixed costs).

@@

```{r Figure_7, fig.cap="Mismatches. X axis labels indicate the model used to derive the policy, followed by the model used to simulate the penalty. Blue bars show the net present value of that scenario.  Compare to red bars, which show the consequence of using a policy that ignores costs all together in the same scenario (i.e. costs as indicated by second code), rather than assuming the policy indicated by the first code in the pair.  All values are scaled to a fraction of the adjustment free net present value.  Successive panels show increasing reduction of the net present value by increasing penalty coefficients: e.g. a policy in the first panel should lose only 10% of its net present value to adjustment costs if it had the correct form of those costs.", dev=c("pdf", "svg", "png")}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

@@

Standing in sharp contrast to these smoothing predictions is our
finding that including policy adjustment costs can actually _increase_ the
variability of quotas through time if there is a fixed cost associated
with making any changes to current policy (e.g. costs of running relevant
stakeholder meetings and public consultations on proposed policy changes,
$\Pi_3$). With this formulation, as stock sizes vary in response to
recruitment, the fishery manager must balance the cost of brokering
a policy change with the cost in forgone fisheries revenue from not
responding to favorable recruitment pulses. The optimal policy involves
ignoring small variations in recruitment, but then assigns a larger quota
when particularly strong recruitment years arise than would have been
the case in the absence of policy adjustment costs. The other functional
form we consider, in which the costs of policy adjustment scale linearly
with the size of the adjustment ($Pi_1$), has less obvious effects on
optimal policies. But interestingly, it is the only formulation for
which we observed positive unchanging quotas of the type most commonly
encountered in real-world applications.

@@

@@

@@

```{r table-caption}
caption = "Table 1: Column 'ignoring' shows the results of ignoring a adjustment penalty when it is present, Column 'assuming' shows the results of assuming an adjustment cost when it is absent. Values (ignoring, assuming) are given as the expected net present value (ENPV) as a fraction of that when no penalty is present, ENPV0.  Percent reduction refers to penalty coefficient being calibrated to achieve the given percent reduction relative to ENPV0 when the penalty is optimally accounted for.  Comparing the second and third columns shows that ignoring a real cost is a significantly graver error, particularly in the case of quadratic costs."
```

```{r Table1, results="asis"}
pandoc.table(table1, caption=caption)
```

>>>>

---
title: Optimal management of a stochastically varying population when policy adjustment is costly
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTK
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul Armsworth 
    affiliation: UTK

@@

options(digits=2)

@@

```{r Figure_1, fig.cap="ICCAT Stock assessments and US Quotas for Bluefin Tuna, 1987 - 2007", fig.width=6, fig.height=4, dev=c("pdf", "svg", "png")}
tuna <- read.csv("components/data/tuna.csv")
tuna<-melt(tuna, id="year")
ggplot(tuna, aes(year, value)) + geom_point() +
  facet_wrap(~variable, ncol=1, scale="free_y") + 
  ylab("Bluefin Tuna harvest (tonnes)") + theme_bw()
```

@@

As the example in Figure 1 makes clear, management policies will rarely be
as responsive as this constant escapement policy assumes. In this paper,
we explore conditions under which this lack of responsiveness may be a
rational response on the part of managers. Specifically, we consider a
case where managers seek to balance the benefits in terms of increased
profits from fishing from more finely tracking recruitment variations
with the growing costs associated with adjusting policies quickly to do
so. The policy adjustment costs involved could reflect pure administrative
transaction costs or preferences held by fishermen, fish processing plants
or other stakeholders, for less variable quotas. In seeking to account
for these policy adjustment costs, we recognize that we do not know just
what functional form they should take and that it will require a lot of
empirical work to estimate that. Therefore, we scope three candidate
functional forms that represent different assumptions about how these
costs operate to determine whether the results we obtain are sensitive
to such differences.

@@

```{r Figure_2, fig.cap="Expected net present value by functional form of penalty. Horizontal axis shows the coefficient $c_i$ governing the magnitude of the policy cost, while vertical axis shows fraction of the maximum expected net present value dissipated by the policy cost, $\\tfrac{NPV_0 - NPV_i}{NPV_0}$.  The horizontal line indicates a value of the stock that is reduced by 25% from the maximum expected value in the absence of policy adjustment costs, $NPV_0({\\bf h_0^*})$. Selecting the coefficient $c_i$ corresponding to this value in each functional form allows us to make consistent comparisons across the different functional forms of policy costs.", dependson="quadcosts", fig.width=5, fig.height=3, cache = fig.cache, dev=c("pdf", "svg", "png")}

@@

ggplot(fees, aes(c2, (npv0-value)/npv0, lty=variable, color=variable)) + 
  geom_line() + 
  geom_hline(aes(yintercept=reduction), linetype=4) + 
  xlab("Penalty coefficient") +
  ylab("Reduction in net present value") + 
  scale_linetype_discrete(labels=relabel) + 
  scale_color_discrete(labels=relabel)

@@

```{r Figure_3, dependson="tidy", fig.cap="Example realization of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=5, fig.width=8, cache = fig.cache, dev=c("pdf", "svg", "png")}

@@

The comparisons shown in Figure 3 are for one realization (i.e. a
particular sequence of random number draws representing environmental
variability) and are made for a particular magnitude of policy
adjustment costs equal to `r reduction *100`% of the maximum expected net present value 
in the base case ($NPV_0({\bf h_0^*})$). Figure 4 generalizes to show
how optimal harvest levels and corresponding optimal stock sizes are
affected by the magnitude of the costs of policy adjustment. This time
we show summary statistics across 50 replicate sets of environmental
conditions. We show the impact of policy adjustment costs on the
variance of both harvests and stock sizes through time; on the
autocorrelation of harvests and stock sizes through time.

```{r Figure_4, fig.cap="Variance and autocorrelation of stock dynamics and harvest dynamics, as a function of the penalty coefficient.  Adjustment penalty is calculated as percentage of the maximum expected NPV in the basic case with no policy adjustment costs.", fig.height=8, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png")}

library("dplyr")

ggplot(filter(stats_df, variable != 'cross.correlation'), 
       aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn))  +
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  coord_cartesian(xlim = c(0, .3)) + 
  facet_wrap(~ variable, scale="free_y") + 
  scale_color_discrete(labels=relabel, name="Penalty function") + 
  scale_fill_discrete(labels=relabel, name="Penalty function") +
  xlab("Penalty size (as fraction of NPV0)")

write.csv(stats_df, "components/data/figure4.csv")
```

For example, when small policy adjustment costs cost little but large
adjustments are expensive ($\Pi_2$), we see the smoothing signal that we
might have expected (see also @Ludwig1980 for this particular case). As
cost penalties increase in severity (i.e., $c_i$ increases), the
variance in quotas through time decreases and the autocorrelation of
quotas through time increases. This smoothing of the harvest quotas
through time has knock-on effects for the corresponding optimal stock
sizes, which also become more autocorrelated as the scaling on policy
adjustment costs ($c_i$) is increased. The autocorrelation in stock
sizes here reflects the lag involved in a more modest increase in
harvests taking longer to eliminate the signal of strong recruitment
events, when compared to a more immediately responsive constant
escapement policy.

@@

Finally, the case where policy adjustment costs scale linearly with the
size of the adjustment ($\Pi_1$ in Eqn. 4) appear to be something of a
middle of the road strategy, in that increasing the severity of policy
adjustment costs (increasing $c_i$ has little effect on the variance or
autocorrelation of optimal harvest rates or stock sizes. To reveal the
particular impact of policy adjustment costs of this type requires a
more targeted summary statistic. Specifically, for each run we
calculated the frequency with which the optimal policy involved
maintaining a positive quota across multiple time steps unaltered. This
type of policy is arguably the most commonly observed behavior in TAC
management, but is one that is almost never observed to be part of the
optimal management strategy in the basic model without policy adjustment
costs (Eqn. 3) or when optimizing against $\Pi_2$ or $\Pi_3$.

@@

While still not common for the particular parameter combinations we examine,
we find that positive unaltered quotas through time are much more likely
to occur when optimizing against $\Pi_1$ where policy adjustment costs
scale linearly with the size of proposed policy changes: over the 100 replicates
time series the harvest policy is strictly positive and identical in consecutive intervals
only `r no_shifts[["L2"]] * 100`% of the time for quadratic costs $\Pi_2$, compared with
`r no_shifts[["L1"]] * 100`% with linear costs $\Pi_1$ and `r no_shifts[["fixed"]]*100`%
for fixed costs $\Pi_3$. Moreover, these occurrences increase in frequency as the severity of
these costs ($c_i$) increases.

@@

```{r Figure_5, fig.cap="Distribution of profits from fishing and costs paid to adjustment.  Distributions are shown when managing a system in which the adjustment costs present are determined by the model indicated, and compared to the same costs and profits of managing the same system using a Reed model, which ignores adjustment costs, and thus tends to pay much higher fees for adjustment.  In contrast, the difference in profit after accounting for costs is only marginally less than the value derived by managing the population to optimize profits when ignoring adjustment costs (Reed model).", fig.width=8, fig.height=6, dependson="histogram-calc", dev=c("pdf", "svg", "png")}

@@

```{r Figure_6, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue), shown by penalty function over increasing magnitudes of cost.  All values are relative to the cost-free adjustment optimum, NPV0.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. The severity of the latter mistake increases sharply as adjustment costs make a larger fraction of the total costs.", dev=c("pdf", "svg", "png"), fig.width=7}

@@

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure 7 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10%
less than optimal free based on the $c_i$ calibration. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

For fixed costs ($\Pi_3$), the reverse is true.  Because fixed costs increase the
volatility of the optimal solution while $\Pi_2$ costs decrease it, simply
ignoring the cost of adjustment all together results in a higher value
than either assuming costs are quadratic ($\Pi_2$) when they are fixed ($\Pi_3$),
or assuming they are fixed when they are quadratic.  Scenarios in which true costs are quadratic (regardless of what
cost structure is assumed by the policy) do significantly worse than
those that are not (i.e. all cases that assume quadratic but actually
have linear or fixed costs).

@@

```{r Figure_7, fig.cap="Mismatches. X axis labels indicate the model used to derive the policy, followed by the model used to simulate the penalty. Blue bars show the net present value of that scenario.  Compare to red bars, which show the consequence of using a policy that ignores costs all together in the same scenario (i.e. costs as indicated by second code), rather than assuming the policy indicated by the first code in the pair.  All values are scaled to a fraction of the adjustment free net present value.  Successive panels show increasing reduction of the net present value by increasing penalty coefficients: e.g. a policy in the first panel would lose only 10% of its net present value to adjustment costs if it had the correct form of those costs, as determined by the choice of penalty coefficient.", dev=c("pdf", "svg", "png"), fig.width=7}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

@@

Standing in sharp contrast to these smoothing predictions is our
finding that including policy adjustment costs can actually _increase_ the
variability of quotas through time if there is a fixed cost associated
with making any changes to current policy (e.g. costs of running relevant
stakeholder meetings and public consultations on proposed policy changes,
$\Pi_3$). With this formulation, as stock sizes vary in response to
recruitment, the fishery manager must balance the cost of brokering
a policy change with the cost in forgone fisheries revenue from not
responding to favorable recruitment pulses. The optimal policy involves
ignoring small variations in recruitment, but then assigns a larger quota
when particularly strong recruitment years arise than would have been
the case in the absence of policy adjustment costs. The other functional
form we consider, in which the costs of policy adjustment scale linearly
with the size of the adjustment ($\Pi_1$), has less obvious effects on
optimal policies. But interestingly, it is the only formulation that
most frequently produces stretches of strictly positive but unchanging
quotas of the type most commonly encountered in real-world applications.

@@

Acknowledgements
================

This project arose out of a NIMBioS working group, "Pretty Darn Good Control".
The authors acknowledge helpful discussions and input from working group
co-organizers Megan Donahue and Carl Toews, and participants Marie-Josee Fortin, 
Dan Ryan, Frank Doyle, Claire Paris, Iadine Chades and Mandy Karnauskas.   
We also acknowledge the following support: NSF Grant DBI-1306697 (CB), ...

@@

_Economic penalties promised for the supplement are still to come_

@@

<!-- Ingoring this, not useful-->
```{r Table1, results="asis", eval=FALSE}
caption = "Table 1: Column 'ignoring' shows the results of ignoring a adjustment penalty when it is present, Column 'assuming' shows the results of assuming an adjustment cost when it is absent. Values (ignoring, assuming) are given as the expected net present value (ENPV) as a fraction of that when no penalty is present, ENPV0.  Percent reduction refers to penalty coefficient being calibrated to achieve the given percent reduction relative to ENPV0 when the penalty is optimally accounted for.  Comparing the second and third columns shows that ignoring a real cost is a significantly graver error, particularly in the case of quadratic costs."
pandoc.table(table1, caption=caption)
```

### 2014-07-25 manuscripts/presentation.Rmd
updated presentation, minor tweaks to manuscript from Paul feedback

---
title: "Fisheries management when quotas are costly to change"
date: "08/12/2014"
output: 
  ioslides_presentation:
    self_contained: false
    css: components/style.css
    incremental: true

author: |
        Carl Boettiger, Michael Bode, James Sanchirico, 
        Jacob LaRiviera, Alan Hastings, Paul Armsworth

@@

<img src="components/images/global-change.jpg" width="500" style="margin:0px auto;display:block">

@@

# Why?

## Postulate: policy intertia

@@

## Dynamic optimization - Dynamic policies

@@

theme_set(theme_tufte())
```
```{r analysis, include=FALSE}
source("components/analysis.R")
```

## Bluefin Tuna Quotas, ICCAT

```{r Figure_1, fig.cap="Quota on Blue Fin Tuna Catch", dev=c("svg", "png")}
tuna <- read.csv("components/data/tuna.csv")
ggplot(tuna, aes(year, quota)) + geom_point() + theme_bw()
```

@@

```{r Figure_2, fig.cap="",  dev=c("svg", "png")}

@@

```{r Figure_3, dependson="tidy", fig.cap="", dev=c("svg", "png")}

@@

```{r Figure_4, fig.cap="", dev=c("svg", "png")}
ggplot(subset(figure4_df, statistic != "cross.correlation"), 
       aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn))  +
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  coord_cartesian(xlim = c(0, .3)) + 
  facet_wrap(~ timeseries + statistic, scale="free_y") + 
  scale_color_discrete(labels=relabel, name="Penalty function") + 
  scale_fill_discrete(labels=relabel, name="Penalty function") + 
  xlab("Penalty size (as fraction of NPV0)")

@@

```{r Figure_5, fig.cap="",  dependson="histogram-calc", dev=c("svg", "png")}

@@

```{r Figure_6, fig.cap="", dev=c("svg", "png")}

@@

```{r Figure_7, fig.cap="", dev=c("svg", "png")}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
```

>>>>

---
title: "Fisheries management when quotas are costly to change"
date: "08/12/2014"
output: 
  ioslides_presentation:
    self_contained: false
    css: components/style.css
    incremental: true
transition: none
author: |
        Carl Boettiger, Michael Bode, James Sanchirico, 
        Jacob LaRiviera, Alan Hastings, Paul Armsworth

@@

<img src="components/images/global-change.jpg" width="500">

@@

theme_set(theme_tufte(base_size=18))
```
```{r analysis, include=FALSE}
source("components/analysis.R")
```

----

![](components/images/tuna_cc0.jpg)

<small> NOAA </small>

## Bluefin Tuna Stocks vs US Quota, ICCAT 1987 - 2007

```{r Figure_1, fig.cap="Quota on Blue Fin Tuna Catch", dev=c("svg", "png"), cache=FALSE}
tuna <- read.csv("components/data/tuna.csv")
tuna<-melt(tuna, id="year")
ggplot(tuna, aes(year, value)) + geom_point() +
  facet_wrap(~variable, ncol=1, scale="free_y") + 
  ylab("Bluefin Tuna harvest (tonnes)") + theme_bw()

## dynamic optimization, dynamic policies

<img src="components/images/mangel_and_clark.jpg", width="300">

----

<img src="components/images/confusion0.png" width="600">

----

<img src="components/images/confusion1.png" width="600">

----

<img src="components/images/confusion_matrix.png" width="600">

----

<img src="components/images/numbers.png" width="600">

----

<img src="components/images/outcomes_numbers.png" width="600">

----

<img src="components/images/time_grid.png" width="600">

----

<img src="components/images/previous_action.png" width="600">

<!--
## Postulate: policy intertia

@@

----

@@

-->

@@

```{r Figure_2, fig.cap="",  dev=c("svg", "png"), cache=FALSE}

@@

```{r Figure_3, dependson="tidy", fig.cap="", dev=c("svg", "png"), cache=FALSE}

@@

```{r Figure_4, fig.cap="", dev=c("svg", "png"), cache=FALSE}
ggplot(subset(figure4_df, statistic != "cross.correlation"), 
       aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn))  +
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  coord_cartesian(xlim = c(0, .3)) + 
  facet_wrap(~ timeseries + statistic, scale="free_y") + 
  scale_color_discrete(labels=relabel, name="Penalty function") + 
  scale_fill_discrete(labels=relabel, name="Penalty function") + 
  xlab("Penalty size (as fraction of NPV0)")

@@

```{r Figure_5, fig.cap="",  dependson="histogram-calc", dev=c("svg", "png"), cache=FALSE}

@@

```{r Figure_6, fig.cap="", dev=c("svg", "png"), cache=FALSE}

@@

```{r Figure_7, fig.cap="", dev=c("svg", "png"), cache=FALSE}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function") + theme_tufte(base_size=14)
```

```{r bluefin_landings, dev=c("svg", "png"), eval=FALSE}
landings <- data.table(read.csv("components/data/iccat.csv"))
bluefin <- landings[Species == "BFT" & PartyName=="UNITED STATES" & YearC %in% 1987:2007,]
ggplot(bluefin, aes(x=YearC,y= Qty_t)) + 
  stat_summary(fun.y = "sum", geom="line") + 
  xlab("Year")+ylab("Catch (tonnes)")

#library("rfisheries")
#landings <- of_landings(species="BFT")
#landings <- landings[landings$year %in% 1982:2007,]
#ggplot(landings, aes(year, catch)) + geom_line() + ylab("Global landings (tonnes)")
```
